detect patches

### 2024-02-06 scripts/classify_patches.Rmd
working on fn to detect marginal populations

@@

  # write out summary values of the patch matching
  data.frame(
    noPres = , 
    PresTopSuitability =, # the number of presences located in top suitability patches
    UniquePatchID = , # the number of uniquelly identified patches
    NoPatchesTopSuitability =  # the total number of most suitable patches from the RASTER
  )

  occ_pops <- dplyr::distinct(occ_pops, PatchID, .keep_all = TRUE) |> # this is multiple pts per patch - not necessary for us. 
    sf::st_drop_geometry() |>
    dplyr::select(date, PatchID) |>
    dplyr::arrange(PatchID) |>
    write.csv(row.names = F,
            file.path('../results/occupiedPatches-topSuitability', paste0(r_taxon, '.csv')))

  # for points which were not patched to a most suitable patch, put them into a hydrologic basin

patchTaggR(f[1])
```

>>>>

hu12 <- rast('../../Geospatial_data/WBD_rast/hu12R-line.tif')
plot(hu12)

@@

  # write out summary values of the patch matching

  fp <- '../results/occupiedPatches'
  data.frame(
    noPres = nrow(occs), 
    PresTopSuitability = nrow(occ_pops), # the number of presences located in top suitability patches
    UniquePatchID = nrow(dplyr::distinct(occ_pops, PatchID)), # the number of uniquelly identified patches
    NoPatchesTopSuitability = terra::minmax(r)[2] # the total number of most suitable patches from the RASTER
  ) |>
    write.csv(row.names = F, 
              file.path(fp, paste0(r_taxon, '-summary', '.csv')))

  dplyr::distinct(occ_pops, PatchID, .keep_all = TRUE) |> # this is multiple pts per patch - not necessary for us. 
    sf::st_drop_geometry() |>
    dplyr::select(date, PatchID) |>
    dplyr::arrange(PatchID) |>
    write.csv(row.names = F,
            file.path(fp, paste0(r_taxon, '-topSuitability', '.csv')))

  # for points which were not patched to a most suitable patch, put them into a hydrologic basin
  marginal_pops <- sf::st_as_sf(occs)|>
    dplyr::filter(!ID %in% occ_pops$ID)

  # identify the drainages within the HU12 database
  drain_nos <- terra::extract(r_drain, marginal_pops, bind = TRUE) |>
    sf::st_as_sf()
  r_drain_sub <- terra::mask()

  # mask original suitability map to these areas, and filter >60% suitable habitat
  f1 <- list.files('suitability_maps')

  suitable <- terra::mask(suitable, r_drain_sub)
  msk <- ifel(suitable > 55 & suitable < 80, NA, 1) # let's make sure these areas do not include existing clusters. 
  suitable <- mask(suitable, msk) # this is now a second raster for occupied marginal populations

  writeRaster(paste0('../results/marginal_habitat', r_taxon, '.tif'))
  return(marginal_pops)
}

ob <- patchTaggR(f[1])
```
